---
title: "COVID19 cases"
author: "LPapaiMD"
date: '2020 04 17 '
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## -DATA-
library(data.table)
library(magrittr)
library(htmltools)
library(ggplot2)
library(gganimate)
library(plotly)
library(forecast)

#FUNCTION: Number formatting and rounding
formatNum<-function(number,precision=2){
    format(round(as.numeric(number),precision),nsmall=precision, big.mark = " ")
}
#Replace vectors
rc<-c("cases"="Cases","deaths"="Deaths","ratioDC"="Deaths/Cases ratio","ratioCP"="Cases/1000people ratio")

###################
# Getting the dataset
url <- "https://opendata.ecdc.europa.eu/covid19/casedistribution/csv"
data<-read.csv(url, na.strings = "", fileEncoding = "UTF-8-BOM")

# Converting to data.table
data<-data.table(data)

# specifying date class
data$dateRep<-as.Date(data$dateRep, tryFormats = c("%Y-%m-%d", "%d/%m/%Y"))
setnames(data,"countriesAndTerritories","country")

# Growth Rate time frame
grtu <- 3
setkey(data,dateRep)
# Computing cumulative sums and ratios:
d<-data[,':='(CCases=cumsum(cases),CDeaths=cumsum(deaths),
                            pop=popData2018[1]),
        by=country][CCases>10,':='(GrowthRate= (CCases-shift(CCases,n=grtu))/(shift(CCases,n=grtu)-shift(CCases,n=2*grtu)),
                                        Cumulative.RatioDC=100*CDeaths/CCases,
                                        Cumulative.RatioCP=1000*CCases/popData2018[1]), 
                    by=country][,':='(countryterritoryCode=NULL,geoId=NULL)]
```

## Basics

- Over **2 million** people infected so far.
- ~ 140.000 confirmed **deaths**
- 213 affected **countries**
- ~20% of cases need **in-hospital care**

## Most cases relative to the population

```{r, echo=T, eval=F}
mostcases <- 
  d[dateRep==max(dateRep) & CCases>6,][
    order(-CCases),head(.SD,15)][
      ,.(country,CCases,pop)]
head(mostcases)
```

```{r, echo=F, eval=T}
mostcases<-d[dateRep==max(dateRep) & CCases>6,][
      order(-CCases),head(.SD,15)][,.(country,CCases,pop)]
head(data.frame(Country=mostcases$country,"Cumulative Cases"=formatNum(mostcases$CCases,0),
                Population=formatNum(mostcases$pop,0)))
```

## Countries with exponential growth

```{r, warning=F}
topfive<-d[country %in% mostcases$country,][order(dateRep),]
p<-ggplot(topfive,aes(x=CCases, y=ma(cases,3))) +
  geom_point(aes(size=pop), color="red") + 
  geom_text(aes(label=country), color="black") +
  scale_x_continuous(trans='log2') +
  scale_y_continuous(trans='log2') +
  ggtitle("Trajectory of COVID-19 Confirmed Cases") +
  xlab("Cumulative case count") +
  ylab("New case count") +
  theme(legend.position="none") +
  transition_time(dateRep) +
  labs(subtitle="Date: {frame_time}") +
  shadow_mark(alpha=0.6, size=0.5, color="grey", exclude_layer = 2)
animate(p, fps=15,duration=20)
```

## Shiny app
 - **Map**
 - **Trajectory:** effective preventive measures?
    - the no. of new cases fall in several countries
    - *like Korea, China, Japan, Switzerland, Norway, Australia, Taiwan...*
 - **Deaths/Cases:** effective healthcare system?
    - healthcare systems that experienced a large exponential growth, were overloaded with cases
    - → increased deaths/cases rate